Decade of Lost Precision: A General Framework For Image Feature Matching Without Geometric Constraints

### 2014-10-14 paper.tex
Inserted missing `of`
Reduced strength of second assumption

\begin{enumerate}
    \item{For a query and \emph{target image} we assume that for any point in 
        the \emph{query image} there is at most one real correspondence in the 
    \emph{target image} and no real correspondences in the \emph{query image} itself 
as assumed by \emph{Ratio-Match}.}
    \item{We assume that distances of feature descriptors are well 
            behaved.  More precisely, given $f_q$, a feature from the 
            \emph{query image}, and $f_{match}$, the real correspondence 
            from the set of features in the \emph{target image} 
            $F_{t}$, we assume $\forall f_i \in F_{proposal}, 
            f_i \neq f_{match}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{q}$ with a true 
             correspondence in the \emph{target image}, we assume that 
             the distribution of \emph{uniqueness ratios} using $F_{t}$ as 
             the baseline set is similar to using $F_{q}$, since the 
         two images in the case of true correspondences are bound 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we prove that \emph{Ratio-Match-Ext} is equal or better than \emph{Ratio-Match} in terms both precision and recall.  Consider the nearest neighbor $f_{p}$ of a query feature $f_{q}$ and the two cases where we either have the $f_{p} \in F_{q}$ or $f_{p} \in F_{t}$. For $f_{p} \in F_{t}$ the uniqueness ratio of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{q} \cup F_{t}, F_{b})\\
        &= \text{r}(f_{q}, F_{t}, F_{b})
\end{align*}
Since $F_{p} = F_{t}$ for \emph{Ratio-Match} this shows that when the nearest neighbor is found in the \emph{target image}, the two algorithms behave identically. For $f_{p} \in F_{q}$ \emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{t}, F_{t}).
\end{align*}

@@

According to the second assumption, feature descriptors behave such that when matched, a true correspondence to a query feature will always be the nearest neighbor. For practical implementations of feature detectors and descriptors, this property strongly depends on the differences between \emph{query} and \emph{target image}, as shown in  \cite{mikolajczyk2005performance,tuytelaars2008local}. In the degenerate case where the closest match is not an actual correspondence, there is no reason why this match should be disproportionately closer to the \emph{query feature} than any matches in the \emph{target image}, so in most cases we would expect the \emph{uniqueness ratio} to be high (signifying a bad match). This means that with a lenient threshold (close to 1), the guarantees about precision and recall equivalence between the different ratio algorithms might no longer hold due to the inconsistent descriptor behavior. It is worth noting that if assumption \#2 holds, matching using only nearest neighbor is not necessarily optimal, unless there is a complete scene overlap between the images matched. For cases where there is little or no overlap, a nearest neighbor would indiscriminately return false matches even with well-behaved feature descriptors.

The third assumption states that the \emph{uniqueness ratios} using either the \emph{query} or \emph{target image} as the baseline set would be similar in distribution given that the feature point we are matching has a true correspondence. We can support this assumption by looking at the \emph{uniqueness ratios} returned by \emph{Self-Match} and \emph{Ratio-Match}, since they use the \emph{query} and \emph{target image} respectively to calculate the \emph{uniqueness ratio}.  Figure~\ref{fig:ratio_hist} shows the actual \emph{uniqueness ratios} measured from 3024 image pairs featuring 3D objects (see below for more on this dataset).  For ratios lower than $0.7$, the \emph{uniqueness ratios} are similar, but as we approach more lenient thresholds, the ratios based on the \emph{query image} are higher than those for the \emph{target image}. This means that the third assumption is invalid for more lenient thresholds, and we can no longer expect \emph{Mirror-Match} to outperform \emph{Ratio-Match-Ext} in terms of precision without a penalty in recall.

>>>>

\begin{enumerate}
    \item{For a query and \emph{target image} we assume that for any point in 
        the \emph{query image} there is at most one real correspondence in the 
    \emph{target image} and no real correspondences in the \emph{query image} itself 
as assumed by \emph{Ratio-Match}.}
    \item{We assume that the distance between two feature descriptors within the \emph{query image} is larger than their distance to a real correspondence in the \emph{target image}.  More precisely, given $f_q$, a feature from the \emph{query image} for which a real correspondence, $f_{match}$, exists in the \emph{target image}, we assume that $\forall f_i \in F_{q}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{q}$ with a true 
             correspondence in the \emph{target image}, we assume that 
             the distribution of \emph{uniqueness ratios} using $F_{t}$ as 
             the baseline set is similar to using $F_{q}$, since the 
         two images in the case of true correspondences are bound 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we prove that \emph{Ratio-Match-Ext} is equal
or better than \emph{Ratio-Match} in terms of both precision and recall.  Consider the nearest neighbor $f_{p}$ of a query feature $f_{q}$ and the two cases where we either have the $f_{p} \in F_{q}$ or $f_{p} \in F_{t}$. For $f_{p} \in F_{t}$ the uniqueness ratio of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{q} \cup F_{t}, F_{b})\\
        &= \text{r}(f_{q}, F_{t}, F_{b})
\end{align*}
Since $F_{p} = F_{t}$ for \emph{Ratio-Match} this shows that when the nearest neighbor is found in the \emph{target image}, the two algorithms behave identically. For $f_{p} \in F_{q}$ \emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{t}, F_{t}).
\end{align*}

@@

According to the second assumption, feature descriptors behave such that when matched, a true correspondence to a \emph{query feature} will always be closer in distance than any other feature in the query image. For this assumption to be violated either the \emph{query image} contains repeating patterns (violating assumption \#1) or the two images are different enough (i.e.\ taken from very different viewpoints or under different lighting conditions) that the descriptor distance between two unrelated features ends up being closer than the distance between a feature and its true correspondence. In this case \emph{Ratio-Match-Ext} and \emph{Mirror-Match} would discard the match while \emph{Ratio-Match} would still attempt to match the descriptor to a feature in the \emph{target image}. Figure~\ref{fig:no_query} shows how the precision of the nearest neighbor match varies with the ratio of the distance to the closest \emph{target feature} over distance to the closest \emph{query feature}. When this ratio is more than 1 there exists a \emph{query feature} which is better matched than the closest \emph{target feature}. The statistic is gathered over 3024 image pairs featuring 3D objects (see Section~\ref{S:3dobjects}) and the results demonstrate that for the cases where the second assumption is violated and \emph{Ratio-Match} might be advantageous, the nearest neighbor in the target image is less likely to be a correct match. % TODO: Create plot and write more precisely

The third assumption states that the \emph{uniqueness ratios} using either the \emph{query} or \emph{target image} as the baseline set would be similar in distribution given that the feature point we are matching has a true correspondence. We can support this assumption by looking at the \emph{uniqueness ratios} returned by \emph{Self-Match} and \emph{Ratio-Match}, since they use the \emph{query} and \emph{target image} respectively to calculate the \emph{uniqueness ratio}. Figure~\ref{fig:ratio_hist} shows the actual \emph{uniqueness ratios} measured from 3024 image pairs featuring 3D objects (see Section~\ref{S:3dobjects} for more on this dataset).  For ratios lower than $0.7$, the \emph{uniqueness ratios} are similar, but as we approach more lenient thresholds, the ratios based on the \emph{query image} are higher than those for the \emph{target image}. This means that the third assumption is invalid for more lenient thresholds, and we can no longer expect \emph{Mirror-Match} to outperform \emph{Ratio-Match-Ext}.
